\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-21 doc/USER_MANUAL/11_graphics.tex
Buildbot ran the code beautifier.
updates manual for better wiki display

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
large output files. \texttt{MOVIE\_VOLUME} files are saved in the
\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
set, otherwise the velocity field is saved. The look of a movie is
determined by the half-duration of the source. The half-duration should
be large enough so that the movie does not contain frequencies that
are not resolved by the mesh, i.e., it should not contain numerical
noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
of each source in the \texttt{CMTSOLUTION} file is replaced by
\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
\]
 \textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
select the appropriate value of 1.1 $\times$ smallest period. As
usual, for a point source one can set \texttt{half duration} in the
\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
in the \texttt{Par\_file} to get the highest frequencies resolved
by the simulation, but for a finite source one would keep all the
\texttt{half durations} as prescribed by the finite source model and
set \texttt{HDUR\_MOVIE} = 0.0.
\end{quote}

@@

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.32\textwidth]{figures/movie_surf_1.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_surf_2.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_surf_3.jpg}
\par\end{centering}

\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
showing movie snapshots of vertical velocity components at different
times.}

\label{fig:movie.surf}
\end{figure}

@@

When running xspecfem3D with the 
\texttt{\small MOVIE\_VOLUME}
flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
specified in the main \texttt{\small Par\_file}, 
e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.
The output is saved by each processor at the time interval specified
by 
\texttt{\small NTSTEP\_BETWEEN\_FRAMES}. 
For all domains,
the velocity field is output to files: 
{\small
\begin{verbatim}
proc??????_velocity_X_it??????.bin
proc??????_velocity_Y_it??????.bin
proc??????_velocity_Z_it??????.bin 
\end{verbatim}
}
For elastic domains, the divergence and curl taken from the velocity field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
as well: 
{\small
\begin{verbatim}
proc??????_div_it??????.bin
proc??????_curl_X_t??????.bin
proc??????_curl_Y_it??????.bin
proc??????_curl_Z_it??????.bin
\end{verbatim}
}
The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and 
\texttt{\small proc??????\_curl\_glob\_it??????.bin}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the 
\texttt{\small SUPPRESS\_UTM\_PROJECTION}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.32\textwidth]{figures/movie_volume_1.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_volume_2.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_volume_3.jpg}
\par\end{centering}

\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
components at different times.}

\label{fig:movie.volume}
\end{figure}

@@

{\footnotesize
\begin{verbatim}
make xcombine_vol_data
\end{verbatim}
}
which will create the executable \texttt{bin/xcombine\_vol\_data}.
To output the usage of this executable, type
{\footnotesize
\begin{verbatim}
./bin/xcombine_vol_data
\end{verbatim}
}
without arguments.

@@

\begin{itemize}
\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
only the bounding box is shown.
\end{itemize}
\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
\textsf{Representation} and select \textsf{wireframe of surface} to
display it.
\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
script (either from the standard output or from \texttt{normal\_plane.txt}
file).
\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
lower and upper limits, or change the color scheme.
\item Now load in the source and receiver location file by \textsf{File}
$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
\item For more information about ParaView, see the ParaView Users Guide
\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
\end{enumerate}
\end{enumerate}
\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[scale=0.6]{figures/3D-S-Kernel.pdf}
\par\end{centering}

\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
an epicentral distance of 176 km from the September 3, 2002, Yorba
Linda earthquake. Lower Panel: Vertical source-receiver cross-section
of the 3D S-wave speed model used for the spectral-element simulations
\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
an epicentral distance of 165 km \citep{LiTr06}.}

\label{figure:P-wave-speed-finite-frequency}
\end{figure}

>>>>

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{Par\_file}. Turning
on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
large output files. \texttt{MOVIE\_VOLUME} files are saved in the
\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
set, otherwise the velocity field is saved. The look of a movie is
determined by the half-duration of the source. The half-duration should
be large enough so that the movie does not contain frequencies that
are not resolved by the mesh, i.e., it should not contain numerical
noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
of each source in the \texttt{CMTSOLUTION} file is replaced by
\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
\]
\end{quote}

\textbf{NOTE:} If \texttt{HDUR\_MOVIE} is set to 0.0, the code will
select the appropriate value of 1.1 $\times$ smallest period. As
usual, for a point source one can set \texttt{half duration} in the
\texttt{CMTSOLUTION} file to be 0.0 and \texttt{HDUR\_MOVIE} = 0.0
in the \texttt{Par\_file} to get the highest frequencies resolved
by the simulation, but for a finite source one would keep all the
\texttt{half durations} as prescribed by the finite source model and
set \texttt{HDUR\_MOVIE} = 0.0.

@@

\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.32\textwidth]{figures/movie_surf_1.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_surf_2.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_surf_3.jpg}
\par
\end{centering}
\caption{Visualization using AVS files created by \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
showing movie snapshots of vertical velocity components at different times.}
\label{fig:movie.surf}
\end{figure}

@@

When running xspecfem3D with the
\texttt{\small MOVIE\_VOLUME}
flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
specified in the main \texttt{\small Par\_file},
e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.
The output is saved by each processor at the time interval specified
by
\texttt{\small NTSTEP\_BETWEEN\_FRAMES}.
For all domains,
the velocity field is output to files:
{\small
\begin{verbatim}
proc??????_velocity_X_it??????.bin
proc??????_velocity_Y_it??????.bin
proc??????_velocity_Z_it??????.bin
\end{verbatim}
}
\noindent
For elastic domains, the divergence and curl taken from the velocity field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
as well:
{\small
\begin{verbatim}
proc??????_div_it??????.bin
proc??????_curl_X_t??????.bin
proc??????_curl_Y_it??????.bin
proc??????_curl_Z_it??????.bin
\end{verbatim}
}
\noindent
The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and
\texttt{\small proc??????\_curl\_glob\_it??????.bin}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the
\texttt{\small SUPPRESS\_UTM\_PROJECTION}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).

\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.32\textwidth]{figures/movie_volume_1.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_volume_2.jpg}
\includegraphics[width=0.32\textwidth]{figures/movie_volume_3.jpg}
\par
\end{centering}
\caption{Paraview visualization using movie volume files (converted by \texttt{xcombine\_vol\_data}
and \texttt{mesh2vtu.pl}) and showing snapshots of vertical velocity
components at different times.}
\label{fig:movie.volume}
\end{figure}

@@

{\footnotesize
\begin{verbatim}
make xcombine_vol_data
\end{verbatim}
}
\noindent
which will create the executable \texttt{bin/xcombine\_vol\_data}.
To output the usage of this executable, type

{\footnotesize
\begin{verbatim}
./bin/xcombine_vol_data
\end{verbatim}
}
\noindent
without arguments.

@@

\begin{itemize}
\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
only the bounding box is shown.
\end{itemize}
\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
\textsf{Representation} and select \textsf{wireframe of surface} to
display it.
\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
script (either from the standard output or from \texttt{normal\_plane.txt}
file).
\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
lower and upper limits, or change the color scheme.
\item Now load in the source and receiver location file by \textsf{File}
$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
\item For more information about ParaView, see the ParaView Users Guide
\urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
\end{enumerate}
\end{enumerate}
%
\begin{figure}[H]
\begin{centering}
\includegraphics[width=6in]{figures/3D-S-Kernel.jpg}
\par
\end{centering}
\caption{(a) Top Panel: Vertical source-receiver cross-section of the S-wave
finite-frequency sensitivity kernel $K_{\beta}$ for station GSC at
an epicentral distance of 176 km from the September 3, 2002, Yorba
Linda earthquake. Lower Panel: Vertical source-receiver cross-section
of the 3D S-wave speed model used for the spectral-element simulations
\citep{KoLiTrSuStSh04}. (b) The same as (a) but for station HEC at
an epicentral distance of 165 km \citep{LiTr06}.}
\label{figure:P-wave-speed-finite-frequency}
\end{figure}
